---
title: "HW01"
author: "Rishab Kulkarni"
date: "1/26/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# i.

Likelihood function:

$\prod_{i=1}^{n}\frac{\lambda^{x_i} e^{-\lambda}}{x_i!}$

Expanding likelihood function:

$\frac{\lambda^{\sum_{i=1}^{n}x_i}e^{-n\lambda}}{\prod_{i=1}^{n}x_i!}$

Log likelihood function:

$\ln\lambda^{\sum_{i=1}^{n}x_i}+\ln e^{-n\lambda}-\ln\prod_{i=1}^{n}x_i!$

Simplifying log likelihood function:

$\sum_{i=1}^{n}x_i\ln\lambda-n\lambda-\ln\prod_{i=1}^{n}x_i!$

Derivative of log likelihood function with respect to $\lambda$:

$\frac{\sum_{i=1}^{n}x_i}{\lambda}-n$

Now, we set this expression equal to 0 and solve for $\lambda$:

$\frac{\sum_{i=1}^{n}x_i}{\lambda}-n = 0$

$\frac{\sum_{i=1}^{n}x_i}{\lambda} = n$

$\frac{\sum_{i=1}^{n}x_i}{n}=\hat{\lambda}$

$\hat{\lambda}=\bar{x}$


# ii.

The exact variance of $\hat{\lambda}$:

The variance of $\hat{\lambda}$ equals the variance of
$\bar{x}$, which is $\frac{\sigma_x^2}{n}$. The variance
of Poisson distribution is $\lambda$, so the exact
variance equals $\frac{\lambda}{n}$.

# iii.

Fisher information of the Poisson distribution equals
$\frac{n}{\lambda}$. The asymptotic variance of $\hat{\lambda}$
is the inverse of the Fisher information. Thus, the
asymptotic variance equals $\frac{\lambda}{n}$, which
is the same as in (ii).

# 2.

MLE estimate for $\hat{\lambda}$ is $\bar{x}$, so we can
use that in our Wald confidence interval.

```{r}
x <- c(9,16,17,14,17,11,12,15,13,9)

x_bar <- mean(x)

# Wald CI
x_bar +c(-1,1) * qnorm(1-0.05/2) * sqrt(x_bar/10)
```

# 3.

Wald test statistic:

$\frac{\hat{\lambda}-\lambda_0}{\sqrt\frac{\lambda}{n}}$

```{r}

# Wald test statistic
w <- (x_bar-10) / sqrt(x_bar/10)
w

# p-value
1 - pnorm(w)
```

# 4.

```{r}

# p-value
1 - ppois(134, 100)
```

# 5.

```{r}

# range
r <- 0:10

# Poisson density function

plot(r, dpois(r, lambda = 6), type='h')
title(main = "Poisson")

# Binomial density function

plot(r, dbinom(r, size= 200, prob= 0.03), type='h')
title(main = "Binomial")
```

Observations:

The density functions of the Poisson and Binomial random variables are
nearly the same. However, P(X = 0), P(X = 1), ... P(X = 5) in the 
Binomial density function are less than P(X = 0), P(X = 1), ... P(X = 5)
in the Poisson density function. P(X = 8), P(X = 9), and P(X = 10) are 
also less in the Binomial density function than P(X = 8), P(X = 9), and 
P(X = 10) in the Poisson density function.

