---
title: "HW03"
author: "Rishab Kulkarni"
date: "2/9/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Question 1**

```{r}
# both conditional odds ratios
(14*381)/(49*75)
(138*416)/(260*134)

# marginal odds ratio
(152*797)/(335*183)
```

**i.**
The conditional odds ratios are 1.451429 for the non-smoker group,
and 1.647761 for the smoker group.

The marginal odds ratio comes from the marginal table, which is
obtained by summing over the levels of the confounding variable, Z
or smoking status. The marginal odds ratio is 1.976087

**ii.**
I think the stratification on smoking status was important in this
study, because it's a confounding variable. Smoking status can change 
the relationship between coffee drinking and MI.

**iii.**
The relative risk cannot be directly estimated because we're dealing
with a retrospective study, which allows only $\hat{\theta}$ to be
estimated. In retrospective studies, we fix the number of outcomes,
so $\hat{\pi_1}$ and $\hat{\pi_2}$ aren't meaningful or significant.

**Question 2**

**i.**
$\pi_{1,2} = 1/14 = 0.071$

**ii.**
$\pi_{1|3} = 7/9 = 0.778$

$\pi_{1|1} = 7/12 = 0.583$


**Question 3**

**i.**
```{r}
# sample odds ratio
((32/36)*(4/36))/((13/16)*(3/16))
```

**ii.**
```{r}
# relative risk
((4/36)/(3/16))
```

**Question 4**

**i.**
```{r}
# odds ratio
(1198*1278)/(1493*567)
```

**ii.**
Yes, these results indicate unfairness. An odds ratio of 1.8 means that
males were ~1.8 times more likely to be admitted at the University of
California, Berkeley than females in 1972.

**iii.**
A useful stratification variable can be the admission major Just like how 
we stratified smoking status into non-smoker and smoker in #1, we can 
stratify academic major into economics, statistics, math, etc. 

Stratifying academic major upon admission can give further insight on
the unfairness associated with admission into certain majors depending
on male or female applicant.

**Question 5**

```{r}
# sensitivity
(41/(41+0))

# specificity
(35/(35+183))
```

The sensitivity and specificity of Dyspnea as a test for CHF are
1 and 0.1605505, respectively.

