---
title: "HW06"
author: "Rishab Kulkarni"
date: "2/28/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**1.**

```{r}
library(pscl)
data("prussian")

# fitting model1, model2

m.1 <- glm(y ~ corp,
           family = poisson,
           prussian)

m.2 <- glm(y ~ corp + year,
           family = poisson,
           prussian)
```

**(i)**
```{r}
library(lmtest)

lrtest(m.1, m.2)
```

p-value from likelihood-ratio test is 0.1305

at the 5% level, we fail to reject the null
hypothesis that death and year are independent.


```{r}
waldtest(m.1, m.2)
```

p-value from Wald test is 0.1324

at the 5% level, we do not reject the null
hypothesis that death and year are independent.

both the likelihood-ratio and Wald test
have the same conclusion. 

we've ample evidence that suggests the
model without the year variable is a
better fit.

**(ii)**

the maximum likelihood estimator for the
poisson distribution is the sample mean
of observations. 

assuming the death rates are poisson, 
the mle of the poisson rate $\lambda$
is the sample mean of the response.

```{r}
mean(prussian$y)
```

**(iii)**

to solve this, we must use the poisson pmf,
so we need lambda.

$\lambda=\mu(x)=e^{\alpha+\beta{x}}$

```{r}
p <- predict.glm(m.1, data.frame(corp="III"), type = "response")

dpois(3, p)
```

**2.**

**(i)**
```{r}
library(faraway)

data("discoveries")

year <- 1860:1959

# poisson model
d <- glm(discoveries ~ year,
         family = poisson,
         discoveries)

waldtest(d)
```

p-value for the Wald test is 0.008059

**(ii)**

```{r}
summary(d)
```

the mean number of discoveries each year 
changes by a factor of $e^{-0.005360}$ or ~0.995

**(iii)**

```{r}
# lower bound
exp(-0.005360 - qnorm(1-0.05/2)*0.001982)

# upper bound
exp(-0.005360 + qnorm(1-0.05/2)*0.001982)
```

**(iv)**

to solve this, we must use the poisson pmf,
so we need lambda.

$\lambda=\mu(x)=e^{\alpha+\beta{x}}$

```{r}
m <- predict.glm(d, data.frame(year=1960), type = "response")

dpois(4, m)
```

**(v)**

```{r}
1-pchisq(deviance(d), df.residual(d))
```

at the 5% level, we reject the null hypothesis
that the model is a good fit of the data.

thus, the deviance doesn't support that the 
model fits the data.

**3.**

**(i)**

```{r}
data("esdcomp")

# fitting model
ratemod <- glm(complaints ~ residency + gender +
                 offset(log(visits)) + I(visits/hours),
               family = poisson, esdcomp)

summary(ratemod)
```

$e^{-0.36733}$ is the average change in mean 
rate of complaints between those who are in 
residency and those who aren't.

$e^{0.07292}$ is the average change in mean
rate of complaints between males and females.

for every one-unit increase in the visits/hr 
rate, the mean rate of complaints changes by
a factor of $e^{0.49344}$

the rate of complaints changes by a factor of 
$e^{-7.33596}$, with other variables set to 0.

**(ii)**

```{r}
newdata = data.frame(gender = "F", residency = "Y",
                     visits = 2500, hours = 1250)

a <- predict.glm(ratemod, newdata, type = "response")

1-ppois(4, a)
```

**4.**

```{r}
# p-value

1-pchisq(deviance(ratemod), df.residual(ratemod))
```

at the 5% level, we reject the null hypothesis
that ratemod is a good fit of the data.

**5.**

**(i)**

```{r}
mean(esdcomp$visits/esdcomp$hours)

sd(esdcomp$visits/esdcomp$hours)
```

**(ii)**

for a k-unit increase, the response changes
by a factor of $e^{\beta{k}}$.

thus, the rate of complaints will change
by a factor of $e^{\beta/4}$.

$e^{\frac{0.49344}{4}}=1.1313$

the rate of complaints changes by ~13%
for every extra 0.25 visits/hr

