---
title: "HW07"
author: "Rishab Kulkarni"
date: "3/9/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**1.**

```{r}
library(tidyverse)
library(faraway)
data("pima")

pima <- pima %>%
  filter(`bmi`!=0.0, `triceps`!=0)

m1 <- glm(test ~ bmi + age,
          family = binomial,
          pima)

m2 <- glm(test ~ bmi + age + 
            triceps + pregnant,
          family = binomial, pima)

summary(m2)
```

The parameters in logistic regression 
are the log odds ratio associated with 
a one-unit increase of the predictor.

For every one-unit increase in bmi, the 
chances of developing diabetes changes 
by a factor of $e^{0.09755}$ or 1.1025

For every one-unit increase in age, the 
chances of developing diabetes changes 
by a factor of $e^{0.04893}$ or 1.05

For every one-unit increase in triceps, 
the chances of developing diabetes changes 
by a factor of $e^{0.01182}$ or 1.012

For every one-unit increase in pregnant, 
the chances of developing diabetes changes 
by a factor of $e^{0.07912}$ or 1.082

**2.**

```{r}

plot(pima$bmi, m2$fitted.values,
     xlab = "bmi", ylab = "fit",
     main = "bmi", col="red")

plot(pima$age, m2$fitted.values,
     xlab = "age", ylab = "fit",
     main = "age", col="blue")

plot(pima$triceps, m2$fitted.values,
     xlab = "triceps", ylab = "fit",
     main = "triceps", col="purple")

plot(pima$pregnant, m2$fitted.values,
     xlab = "pregnant", ylab = "fit",
     main = "pregnant", col="green4")
```


```{r}
par(mfrow=c(2,2))

plot(pima$bmi, predict.glm(m2),
     pch=19, cex = 0.5, col="red",
     xlab = "bmi", ylab = "fit")

plot(pima$age, predict.glm(m2),
     pch=19, cex=0.5, col="blue",
     xlab = "age",ylab = "fit")

plot(pima$triceps, predict.glm(m2),
     pch=19, cex=0.5, col="darkgreen",
     xlab = "triceps", ylab = "fit")

plot(pima$pregnant, predict.glm(m2),
     pch=19, cex=0.5, col="purple",
     xlab = "pregnant", ylab = "fit")

```

```{r}
library(boot)

# leverages
lev <- glm.diag(m2)$h


which.max(lev)

pima %>%
  filter(`age`==28)

```

case 410 had the highest leverage

After examining her data, I found that her
insulin value was much higher than other
cases. Her glucose and triceps were also
faintly higher than other cases.

case 410 had an age of 28 and since age
was a significant predictor, I filtered
the data to include only cases where age
was 28. Then, I compared case 410 to the
other cases.

case 410 had much higher insulin than 
other cases; glucose, triceps, and bmi
were also higher in case 410 than the
other cases.

also compared case 410 to other cases
without filtering for age and case 410
had much higher insulin - 579, than other 
cases in the data. She also had higher 
glucose and triceps.

glucose and triceps were a bit higher,
but insulin was way higher than the
insulin in other cases. Thus, I can
believe the glucose and triceps value,
but not the insulin value.


```{r}
anova(m1, m2, test = "Chisq")
```

p-value was 0.06834

at the 5% level, we fail to reject the
null hypothesis that model1 and model2
fit the data equally well.

we've ample evidence that suggests that
both models fit the data equally well.

**3.**

```{r}
library(lmtest)

lrtest(m1, m2)
```

this time, using `lrtest()`

The p-value was 0.06834 again.

at the 5% level, we fail to reject the 
null hypothesis that the reduced and 
full model fit the data equally well.

we've ample evidence that suggests 
both models fit the data equally well.

**4.**

```{r}
newdata <- data.frame(pregnant=10, bmi=40,
                      age=50, triceps=40)

predict(m2, newdata, type = "response")

predict(m2, newdata, type = "response",
        se.fit = TRUE)$se.fit

```

**5.**

```{r}
# glucose
m.gluc <- glm(test ~ bmi + age + triceps
             + pregnant + glucose,
             family = binomial, pima)

# diastolic
m.dias <- glm(test ~ bmi + age + triceps
             + pregnant + diastolic,
             family = binomial, pima)

# insulin
m.ins <- glm(test ~ bmi + age + triceps
             + pregnant + insulin,
             family = binomial, pima)

# diabetes
m.diab <- glm(test ~ bmi + age + triceps
             + pregnant + diabetes,
             family = binomial, pima)
```


```{r}
anova(m2, m.gluc, test = "Chisq")
```


```{r}
anova(m2, m.dias, test = "Chisq")
```


```{r}
anova(m2, m.ins, test = "Chisq")
```


```{r}
anova(m2, m.diab, test = "Chisq")
```

the glucose predictor resulted in the
smallest p-value of the Wald test for
significance, which was < 2.2e-16.


```{r}
summary(m.gluc)
```

for every one-unit increase in glucose, the
chances of developing diabetes changes by a
factor of $e^{0.031148}$ or ~1.032
