---
title: "HW08"
author: "Rishab Kulkarni"
date: "3/22/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**1i.**

```{r}
data("UCBAdmissions")

berk <- data.frame(UCBAdmissions)

b <- reshape(berk, v.names = "Freq",
             timevar = "Admit",
             idvar = c("Gender", "Dept"),
             direction = "wide")
```

**1ii.**

```{r}
# additive model
add.mod <- glm(cbind(Freq.Admitted, Freq.Rejected) ~ Gender + Dept, family = "binomial", b)

summary(add.mod)
```

After controlling for Dept:

The odds of female admission is
1.105 times the male admissions.

Estimated odds ratio of admission
for females versus males:

```{r}
exp(0.09987)
```

Estimated odds ratio of admission 
to department B verses department A:

```{r}
exp(-0.04340)
```

Estimated odds ratio of admission
to department C versus department A:

```{r}
exp(-1.26260)
```

Estimated odds ratio of admission
to department D verses department A:

```{r}
exp(-1.29461 )
```

Estimated odds ratio of admission
to department E versus department A:

```{r}
exp(-1.73931)
```

Estimated odds ratio of admission
to department F versus department A:

```{r}
exp(-3.30648)
```

**2i.**

```{r}
# residual deviance test p-value

1-pchisq(deviance(add.mod), df.residual(add.mod))
```

**2ii.**

```{r}
# Pearson residuals

resid(add.mod, type = "pearson")

plot(resid(add.mod, type = "pearson"))
```

The outliers seem to be 
indices 3, 9, 17.

These indices correspond to females
admitted to Dept B, males admitted
to Dept C, and males admitted to
Dept E.

**2iii.**

```{r}

x.mod <- glm(cbind(Freq.Admitted, Freq.Rejected) ~ Gender, family = "binomial", b)

summary(x.mod)
```

Note:

Calculated 1/$e^{\hat{\beta}}$
because the summary compares female
admission to male admission; we want
male admission versus female admission.

The odds ratio for male admissions
versus female admission is:

```{r}
1/exp(-0.61035)
```

This odds ratio is higher than the 
same odds ratio of male admission 
versus female admission in the 
additive model.

**2iv.**

```{r}
library(lmtest)

z.mod <- glm(cbind(Freq.Admitted, Freq.Rejected) ~ Dept, family = "binomial", b)

lrtest(add.mod, z.mod)
```

The p-value from the likelihood-ratio 
test was 0.2159

We fail to reject the null hypothesis
that the homogeneous association or
additive model does not hold.

**3i.**

```{r}
library(faraway)
data("wbca")

m <- glm(Class ~., family = "binomial", wbca)
summary(m)
```

At the 5% level - Adhes, BNucl, Chrom,
NNucl, and Thick are significant 
predictors.

**3ii.**

```{r}
# removing USize

m2 <- glm(Class ~. -USize, family = "binomial", wbca)
summary(m2)
```

```{r}
# removing Epith

m3 <- glm(Class ~. -USize-Epith, family = "binomial", wbca)
summary(m3)
```

```{r}
# removing UShap

m4 <- glm(Class ~. -USize-Epith-UShap, family = "binomial", wbca)
summary(m4)
```

```{r}
# removing Mitos

m5 <- glm(Class ~. -USize-Epith-UShap-Mitos, family = "binomial", wbca)

# final model
summary(m5)
```

The standard errors for the coefficients
in the final model are lower than those
in the full model.

**4i.**

```{r}
# backwards elimination

back.m <- step(m)

summary(back.m)
```

Used `step`()

Here, the final model still includes
insignificant predictors at the 5%
level. The final model includes UShap
and Mitos, which are not included in
the final model from part 3 (ii).

**4ii.**

```{r}
# forwards selection

m.int <- glm(Class ~ 1, family = "binomial", wbca)

step(m.int, scope=list(upper=m, lower=~1), direction="forward")
```

```{r}
# summary of final model after forwards selection

fwd.m <- glm(Class ~. -Epith-UShap, family = "binomial", wbca)
summary(fwd.m)
```


The final model using backwards 
elimination was different than 
the final model using forwards 
selection.

The final model using backwards
elimination had an AIC of 105.66,
whereas the final model using
forwards selection had an AIC of
106.9

The final model after using
backwards elimination is the
better fit due to the lower
AIC score.

**5i.**

```{r}
(2^9) - 1
```

**5ii.**

```{r}
n.mod <- choose(9,2) + choose(9,3)*choose(3,2) + choose(9,4)*choose(4,2) + choose(9,5)*choose(5,2) + choose(9,6)*choose(6,2) + choose(9,7)*choose(7,2) + choose(9,8)*choose(8,2) + choose(9,9)*choose(9,2)

# including models w/ no interaction terms
n.mod + 8
```

